Study of Higgs plus single top production using events with a same sign dimuon at the Large Hadron Collider

### 2019-11-12 presentacion_hiram/presentacion-tesis.tex
new

\begin{frame}
\section{Overview and motivation}
\frametitle{Overview}
\begin{itemize}
\item Through this project it was studied the production of Higgs boson in association with a
single top quark ($tH$) in proton-proton collisions. 
This mechanism of production of the Higgs boson has not been observed before by any
experiment.
%\item Standard Model and the Higgs production mechanism
\item A brief introduction to the LHC and CMS
\item Event selection and reconstruction of the tH signal
\item Creation and application of a statistical model for the analysis of events 
\item Making predictions for the future experiments.
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Backgrounds}
\subsection{Backgrounds}
\begin{itemize}
	\item Backgrounds are considered information in the data that is not the signal. 
	\item Backgrounds share certain characteristics of the signal.
	\item Backgrounds are a big factor in the statistical study of the signal.
\end{itemize}
\end{frame}

@@

\begin{minipage}[c]{0.45\textwidth}
\begin{figure}
	\centering
	\includegraphics[scale=0.3]{figures/histo.png}
	\caption*{\tiny{Pre-fit classifier outputs for same sign dimuon, for training against $t\bar{t}V$ }}
\end{figure}
\end{minipage}
\hfill
\begin{minipage}[c]{.45\textwidth}
	\begin{figure}	\centering
	\includegraphics[width=\textwidth]{figures/sc.png}
	\caption*{\tiny{Observed and expected 95$\%$ C.L. upper limit on the tH + t$\bar{t}$H cross section times H $\rightarrow$ WW $+$ $\tau\tau$ $+$ ZZ branching fraction for different 20 values of the coupling ratio $\kappa_t /\kappa_v$ . The expected limit is derived from a background-only MC dataset.}}
	\end{figure}
\end{minipage}
\end{frame}
}

@@

\begin{frame}
\frametitle{Signal and Backgrounds}
\begin{table}
	\centering
	\caption*{{Event yields for signal and backgrounds after the event selection for a integrated luminosity of 35.9 fb$^{-1}$.
			The uncertainties of yields include statistical and systematic \cite{2}}} 
	\begin{tabular}{cc}
		\hline
		Process & Number of events \\
		\hline
		$t\bar{t}W$ & 68 $\pm$ 10 \\
		$t\bar{t}Z$ & 25.9 $\pm$ 3.9\\
		$WZ$ & 15.1$\pm$7.7\\
		Rares & 20.9 $\pm$ 4.9\\
		Fakes & 80.9 $\pm$9.4\\
		$t\bar{t}H$ & 24.2 $\pm$ 2.1 \\
		\hline
		$tH$ (SM) & 2.14 $\pm$ 0.13\\
		$tH$ ($k_t=-1$) &26.2 $\pm$ 2.2
	\end{tabular}	
	\label{tth-table}
\end{table}
\end{frame}

@@

\begin{frame}
\subsection{BDT}
\frametitle{BDT}
\small{
\begin{itemize}
\item A decision tree takes a set of input features and splits input data recursively based on
those features.
Boosting is a method of combining many weak learners (trees) into a strong classifier. The features can be a mix of categorical and continuous data.
\item The BDT training is performed using several event variables for signal and background. 
\end{itemize}
BDT Variables}
\tiny{
\begin{itemize}
\item	Trailing lepton $p_{t}$
\item 	Total charge of tight leptons
\item 	min $\Delta$R (lepton pairs)
\item 	$\Delta\phi$ between highest $p_t$ lepton pair
\item 	Number of jets with |$\eta$| < 2.4
\item	Number of non b-tagged jets with |$\eta$| > 1.0
\item	Maximum |$\eta$| for jets
\item	$\Delta\eta$ (most forward light jet, closest lepton)
\item	$\Delta\eta$ (most forward light jet, hardest loosely b-tagged jet)
\item	$\Delta\eta$ (most forward light jet, 2nd hardest loosely b-tagged jet)
\end{itemize} }
\end{frame}

@@

\begin{frame}
\subsection{Likelihood model}
\frametitle{Likelihood model}
\small{
To estimate the sensibility of the $tH$ signal, we define an Asimov dataset, made by replacing the ensemble of simulated backgrounds and signal by a single one. The statistical uncertainty of the Asimov data is calculated as $\sqrt{n}$, with $n$ the number of events.\\
The likelihood function is the product of Poisson probabilities for all bins
\begin{align}
	L(\mu,\alpha)=\prod_{j=1}^{N}\frac{(\mu s_j +b_j)^{n_j}}{n_j !}e^{-(\mu s_j+b_j)} \prod_{k=1}^M e^{\frac{-\alpha^2_k}{2}}
\end{align}
	where $N$ is the total number of bins, $n_j$ is the number of events in a bin $j$, $s_j$ is the number of signal events, $\mu$ is a parameter that modifies the signal strength and $b_j$ is the number of background events.
	$b_j$ is the sum of different background processes $k$
\begin{align}
b_j=\sum_k^M b_j^k(1+ \alpha_k \sigma_k)
\end{align}
$\alpha_k$ is the parameter that modifies the expected background prediction and $\sigma_k$ is the systematic uncertainty of the associated background. $\sigma_k$ for the backgrounds are shown in the Table of yields. }
\end{frame}

\begin{frame}
\frametitle{Results}
	The fit is applied by minimizing the $-\log{L}$ (NLL) with respect to the parameter $\mu$ and $\alpha_k$. The minimization is performed by using the package ROOFIT. In the next slides, it shows the results of the model fitting using the Asimov data for the SM and the $k_t=-1$.
\end{frame}

@@

\begin{frame}
\frametitle{Results}
\framesubtitle{Likelihood scan}
Due to the large background, the signal strength for the Asimov data with 35.9 fb$^{-1}$ is consistent with zero.
Therefore, we estimate an upper limit on the signal strength at 95$\%$ confidence level.\\

We can define the likelihood ratio
\begin{align}
\lambda(\mu,\alpha)=\frac{L(\mu,\alpha)}{L(\hat{\mu},\hat{\alpha})}
\end{align}
Where $\hat{\alpha}$ and $\hat{\mu}$ are the parameters obtained in the previous section which correspond to the minimal of the NLL.
\end{frame}

@@

\begin{frame}
\frametitle{Results}
\framesubtitle{Upper limit}
To determine an upper limit on the strength parameter $\mu$, we use the following statistical test
\begin{align}
q(\mu,\alpha)= -2\ln{\lambda} 
\end{align}
High values of $q$ represent greater incompatibility between the data and the fit model.
$q$ is a random variable with a $\chi^2$ distribution.
\end{frame}

>>>>

%\begin{frame}
%\section{Overview and motivation}
%\frametitle{Overview}
%\begin{itemize}
%\item Through this project it was studied the production of Higgs boson in association with a
%single top quark ($tH$) in proton-proton collisions. 
%This mechanism of production of the Higgs boson has not been observed before by any
%experiment.
%\item Standard Model and the Higgs production mechanism
%\item A brief introduction to the LHC and CMS
%\item Event selection and reconstruction of the tH signal
%\item Creation and application of a statistical model for the analysis of events 
%\item Making predictions for the future experiments.
%\end{itemize}
%\end{frame}

@@

%\begin{frame}
%\frametitle{Backgrounds}
%\subsection{Backgrounds}
%\begin{itemize}
%	\item Backgrounds are considered information in the data that is not the signal. 
%	\item Backgrounds share certain characteristics of the signal.
%	\item Backgrounds are a big factor in the statistical study of the signal.
%\end{itemize}
%\end{frame}

@@

\begin{minipage}[c]{0.45\textwidth}
\begin{figure}
	\centering
	\includegraphics[scale=0.3]{figures/histo.png}
	\caption*{\tiny{Pre-fit classifier outputs for same sign dimuon, for training against $t\bar{t}V$ }}
\end{figure}
\end{minipage}
\hfill
\begin{minipage}[c]{.45\textwidth}
	\tiny
\begin{table}
	\centering
	\caption*{{Event yields for signal and backgrounds after the event selection for a integrated luminosity of 35.9 fb$^{-1}$.
			The uncertainties of yields include statistical and systematic \cite{2}}} 
	\begin{tabular}{cc}
		\hline
		Process & Number of events \\
		\hline
		$t\bar{t}W$ & 68 $\pm$ 10 \\
		$t\bar{t}Z$ & 25.9 $\pm$ 3.9\\
		$WZ$ & 15.1$\pm$7.7\\
		Rares & 20.9 $\pm$ 4.9\\
		Fakes & 80.9 $\pm$9.4\\
		$t\bar{t}H$ & 24.2 $\pm$ 2.1 \\
		\hline
		$tH$ (SM) & 2.14 $\pm$ 0.13\\
		$tH$ ($k_t=-1$) &26.2 $\pm$ 2.2
	\end{tabular}	
	\label{tth-table}
\end{table}
%	\begin{figure}	\centering
%	\includegraphics[width=\textwidth]{figures/sc.png}
%	\caption*{\tiny{Observed and expected 95$\%$ C.L. upper limit on the tH + t$\bar{t}$H cross section times H $\rightarrow$ WW $+$ $\tau\tau$ $+$ ZZ branching fraction for different 20 values of the coupling ratio $\kappa_t /\kappa_v$ . The expected limit is derived from a background-only MC dataset.}}
%	\end{figure}
\end{minipage}
\end{frame}
}

@@

@@

\begin{frame}
\subsection{BDT}
\frametitle{BDT}
\small{
 A decision tree takes a set of input features and splits input data recursively based on
those features.
Boosting is a method of combining many weak learners (trees) into a strong classifier. The features can be a mix of categorical and continuous data. \\

BDT Variables}
\tiny{
\begin{itemize}
\item	Trailing lepton $p_{t}$
\item 	Total charge of tight leptons
\item 	min $\Delta$R (lepton pairs)
\item 	$\Delta\phi$ between highest $p_t$ lepton pair
\item 	Number of jets with |$\eta$| < 2.4
\item	Number of non b-tagged jets with |$\eta$| > 1.0
\item	Maximum |$\eta$| for jets
\item	$\Delta\eta$ (most forward light jet, closest lepton)
\item	$\Delta\eta$ (most forward light jet, hardest loosely b-tagged jet)
\item	$\Delta\eta$ (most forward light jet, 2nd hardest loosely b-tagged jet)
\end{itemize} }
\end{frame}

@@

\begin{frame}
\subsection{Likelihood model}
\frametitle{Likelihood model}
{\small \begin{itemize}
\item To estimate the sensibility of the $tH$ signal, we define an Asimov dataset, made by replacing the ensemble of simulated backgrounds and signal by a single one. 
\item The statistical uncertainty of the Asimov data is calculated as $\sqrt{n}$, with $n$ the number of events.\\
\item The likelihood function is the product of Poisson probabilities for all bins
\begin{align}
	L(\mu,\alpha)=\prod_{j=1}^{N}\frac{(\mu s_j +b_j)^{n_j}}{n_j !}e^{-(\mu s_j+b_j)} \prod_{k=1}^M e^{\frac{-\alpha^2_k}{2}}
\end{align}
	where $N$ is the total number of bins, $n_j$ is the number of events in a bin $j$, $s_j$ is the number of signal events, $\mu$ is a parameter that modifies the signal strength and $b_j$ is the number of background events.
\end{itemize}}
\end{frame}

\begin{frame}
\frametitle{Likelihood model}
{\small
\begin{itemize}
	\item $b_j$ is the sum of different background processes $k$
	\begin{align}
	b_j=\sum_k^M b_j^k(1+ \alpha_k \sigma_k)
	\end{align}
	$\alpha_k$ is the parameter that modifies the expected background prediction and $\sigma_k$ is the systematic uncertainty of the associated background. $\sigma_k$ for the backgrounds are shown in the Table of yields. 
		\item The fit is applied by minimizing the $-\log{L}$ (NLL) with respect to the parameter $\mu$ and $\alpha_k$. The minimization is performed by using the package ROOFIT. In the next slides, it shows the results of the model fitting using the Asimov data for the SM and the $k_t=-1$.
\end{itemize}}
\end{frame}

@@

\begin{frame}
\frametitle{Results}
\framesubtitle{Likelihood scan}
{\small
	\begin{itemize}
\item Due to the large background, the signal strength for the Asimov data with 35.9 fb$^{-1}$ is consistent with zero.
Therefore, we estimate an upper limit on the signal strength at 95$\%$ confidence level.

\item We can define the likelihood ratio
\begin{align}
\lambda(\mu,\alpha)=\frac{L(\mu,\alpha)}{L(\hat{\mu},\hat{\alpha})}
\end{align}
Where $\hat{\alpha}$ and $\hat{\mu}$ are the parameters obtained in the previous section which correspond to the minimal of the NLL.

\item To determine an upper limit on the strength parameter $\mu$, we use the following statistical test
\begin{align}
q(\mu,\alpha)= -2\ln{\lambda} 
\end{align}
\item High values of $q$ represent greater incompatibility between the data and the fit model.
$q$ is a random variable with a $\chi^2$ distribution.
\end{itemize}
}
\end{frame}

@@
